---
title: "Retail Forecasting Project Model"
author: "Jie Heng Yu"
date: "12/4/2023"
output: html_document
runtime: shiny
---

```{r, include = FALSE}
library(tidyverse)
library(mlr)
library(parallel)
library(parallelMap)
```

# Retail Forecasting Project

### Data Preparation

We'll transform our data so that it only has the variables based off the results of the AIC-selected model from our EDA.

```{r}
retailDF <- as_tibble(read.csv('ForecastingCaseStudy.csv'))
# Rename columns
colnames(retailDF)[4] <- 'PriceDiscount'
colnames(retailDF)[10:12] <- c('Valentines Day', 'Easter', 'Christmas')
# Recode date times to week year
retailDF$date <- as.Date(retailDF$date, '%m/%d/%Y')
retailDF <- mutate(retailDF, Week = isoweek(ymd(retailDF$date)), Year = isoyear(ymd(retailDF$date)))
# Remove % from Price Discount (%)
retailDF$PriceDiscount <- as.numeric(str_remove(retailDF$PriceDiscount, '%'))
# Remove date column
retailDF <- select(retailDF, -date)
# Recode products column to be numeric
retailDF$Product <- recode(retailDF$Product, 'SKU1' = 1, 'SKU2' = 2, 'SKU3' = 3, 'SKU4' = 4, 'SKU5' = 5, 'SKU6' = 6)
colnames(retailDF)[4:6] <- c('In_Store_Promo', 'Catalogue_Promo', 'Store_End_Promo')
colnames(retailDF)[9] <- 'Valentines_Day'
retailDF
```

We'll remove the outliers from our data based on our Sales feature. Outliers in statistical research settings are anything below 25th quartile - 1.5 IQR, or anything above 75th quartile + 1.5 IQR. IQR refers to the inter-quartile range, which is the 75th quartile - 25th quartile. It should be removing large sales numbers only -- can't imagine there being negative sales -- I could be wrong.

```{r}
quant25 <- quantile(retailDF$Sales, c(0, 0.25, 0.5, 0.75, 1))[[2]]
quant75 <- quantile(retailDF$Sales, c(0, 0.25, 0.5, 0.75, 1))[[4]]
retailDF <- filter(retailDF, Sales > (quant25 - 3 * (quant75 - quant25)) & Sales < (quant75 + 3 * (quant75 - quant25)))
retailDF
```

Let's also reduce our dataset to our AIC-selected variables.

```{r}
retailDF <- select(retailDF, Sales, Product, In_Store_Promo, PriceDiscount, Catalogue_Promo, Store_End_Promo, Covid_Flag, Christmas, Week, Year)
retailDF
```

We'll also have to recode our PriceDiscount feature. We'll put the observations into buckets of 10. PriceDiscount will then almost like a polytomous factor in our data set, so we will turn it into a set of k - 1 dummy variables that will contain the same information, all of which are independent but are references of a level of the factor. This will make more sense when I explain it in my presentation. The reason for doing so is to create a model that won't have a unreasonably large MSE, mean squared error, which is a standardized measure of error for regression models.

```{r}
unique(retailDF$PriceDiscount)
retailDF$PriceDiscount <- case_when(retailDF$PriceDiscount %in% 1:10 ~ 1,
                                    retailDF$PriceDiscount %in% 11:20 ~ 2,
                                    retailDF$PriceDiscount %in% 21:30 ~ 3,
                                    retailDF$PriceDiscount %in% 31:40 ~ 4,
                                    retailDF$PriceDiscount %in% 41:50 ~ 5,
                                    retailDF$PriceDiscount %in% 51:60 ~ 6,
                                    retailDF$PriceDiscount %in% 61:70 ~ 7,
                                    retailDF$PriceDiscount %in% 71:80 ~ 8,
                                    retailDF$PriceDiscount %in% 81:90 ~ 9,
                                    retailDF$PriceDiscount %in% 91:100 ~ 10,
                                    TRUE ~ 0)
unique(retailDF$PriceDiscount)
#retailDF$PriceDiscount1 <- case_when(retailDF$PriceDiscount == 1 ~ 1,
#                                     TRUE ~ 0)
#retailDF$PriceDiscount2 <- case_when(retailDF$PriceDiscount == 2 ~ 1,
#                                     TRUE ~ 0)
#retailDF$PriceDiscount3 <- case_when(retailDF$PriceDiscount == 3 ~ 1,
#                                     TRUE ~ 0)
#retailDF$PriceDiscount4 <- case_when(retailDF$PriceDiscount == 4 ~ 1,
#                                     TRUE ~ 0)
#retailDF$PriceDiscount5 <- case_when(retailDF$PriceDiscount == 5 ~ 1,
#                                     TRUE ~ 0)
#retailDF$PriceDiscount6 <- case_when(retailDF$PriceDiscount == 6 ~ 1,
#                                     TRUE ~ 0)
#retailDF$PriceDiscount7 <- case_when(retailDF$PriceDiscount == 7 ~ 1,
#                                     TRUE ~ 0)
#retailDF$PriceDiscount8 <- case_when(retailDF$PriceDiscount == 8 ~ 1,
#                                     TRUE ~ 0)
#retailDF$PriceDiscount9 <- case_when(retailDF$PriceDiscount == 9 ~ 1,
#                                     TRUE ~ 0)
#retailDF <- select(retailDF, -PriceDiscount)
retailDF
```

### Hyperparameter Tuning

We tune our hyperparameters to give us the best possible model given our circumstances.

```{r, message = FALSE, warning = FALSE}
retailTask <- makeRegrTask(data = retailDF, target = 'Sales')
forest <- makeLearner('regr.randomForest')
forestParamSpace <- makeParamSet(
  makeIntegerParam('ntree', lower = 50, upper = 50),
  makeIntegerParam('mtry', lower = 3, upper = 10),
  makeIntegerParam('nodesize', lower = 1, upper = 10),
  makeIntegerParam('maxnodes', lower = 5, upper = 30)
)
randSearch <- makeTuneControlRandom(maxit = 150)
repKFold <- makeResampleDesc('RepCV', folds = 10, reps = 5)

parallelStartSocket(cpus = detectCores() - 1)
tunedForestPars <- tuneParams(forest, task = retailTask, resampling = repKFold, par.set = forestParamSpace, control = randSearch)
parallelStop()
```

### Model Training

Using our tuned hyperparameters, we train our model.

```{r}
tunedForest <- setHyperPars(forest, par.vals = tunedForestPars$x)
tunedForestModel <- train(tunedForest, retailTask)
```

### Prediction Testing

Lets generate some fake data & use it to predict its sales values.

```{r, message = FALSE, warning = FALSE}
newData <- tibble(Product = 1:3,
                  In_Store_Promo = c(0, 1, 0),
                  PriceDiscount = 3:5,
                  Catalogue_Promo = c(1, 0, 1),
                  Store_End_Promo = c(0, 1, 0),
                  Covid_Flag = c(1, 1, 0),
                  Christmas = c(0, 1, 1),
                  Week = 24:26,
                  Year = 2017:2019)
newData
getPredictionResponse(predict(tunedForestModel, newdata = newData))
```

***

# ML Predictions Dashboard

### UI

```{r}
ui <- fluidPage(
  titlePanel('Retail Forecasting Project Predictive Dashboard'),
  sidebarLayout(
    sidebarPanel(
      selectInput(inputId = 'product', label = 'Product:', choices = c('SKU1', 'SKU2', 'SKU3', 'SKU4', 'SKU5', 'SKU6'), selected = 'SKU1'),
      checkboxInput(inputId = 'inStorePromo', label = 'In-Store Promotion', value = FALSE),
      sliderInput(inputId = 'discount', label = 'Discount (%):', min = 0, max = 90, value = 0, step = 1),
      checkboxInput(inputId = 'cataloguePromo', label = 'Catalogue Promotion', value = FALSE),
      checkboxInput(inputId = 'storeEndPromo', label = 'Store Ending Promotion', value = FALSE),
      checkboxInput(inputId = 'covid', label = 'Covid Risk', value = FALSE),
      checkboxInput(inputId = 'christmas', label = 'Christmas', value = FALSE),
      dateInput(inputId = 'date', label = 'Date:', value = '2017-06-30', min = '2017-01-01', max = '2023-12-30', format = 'yyyy-mm-dd')
    ),
    mainPanel(
      verbatimTextOutput(outputId = 'context'),
      verbatimTextOutput(outputId = 'prediction')
    )
  )
)
```

### Server

```{r}
server <- function(input, output, session) {
  output$context <- renderText({'The predicted sales is below.'})
  output$prediction <- renderText({ 
    new_product <- as.numeric(str_remove(input$product, 'SKU'))
    new_store_promo <- case_when(input$inStorePromo == FALSE ~ 0, TRUE ~ 1)
    new_discount <- case_when(input$discount %in% 1:10 ~ 1,
                              input$discount %in% 11:20 ~ 2,
                              input$discount %in% 21:30 ~ 3,
                              input$discount %in% 31:40 ~ 4,
                              input$discount %in% 41:50 ~ 5,
                              input$discount %in% 51:60 ~ 6,
                              input$discount %in% 61:70 ~ 7,
                              input$discount %in% 71:80 ~ 8,
                              input$discount %in% 81:90 ~ 9,
                              TRUE ~ 0)
    new_catalogue_promo <- case_when(input$cataloguePromo == FALSE ~ 0, TRUE ~ 1)
    new_end_promo <- case_when(input$storeEndPromo == FALSE ~ 0, TRUE ~ 1)
    new_covid_risk <- case_when(input$covid == FALSE ~ 0, TRUE ~ 1)
    new_christmas <- case_when(input$christmas == FALSE ~ 0, TRUE ~ 1)
    new_week <- isoweek(ymd(input$date))
    new_year <- isoyear(ymd(input$date))
    new_data <- tibble(Product = new_product,
                       In_Store_Promo = new_store_promo,
                       PriceDiscount = new_discount,
                       Catalogue_Promo = new_catalogue_promo,
                       Store_End_Promo = new_end_promo,
                       Covid_Flag = new_covid_risk,
                       Christmas = new_christmas,
                       Week = new_week,
                       Year = new_year)
    result <- getPredictionResponse(predict(tunedForestModel, newdata = new_data))
    return(result)})
}
```

### Run App

```{r}
shinyApp(ui, server)
```